---
title: "mPGES_11.22.23 GSEA Rds Generation"
author: "Mike Martinez"
date: "2023-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this script is to generate Rds objects of GSEA for the mPGES delta 14 KO comparisons generated in mPGES_11.21.23

```{r set working directory and load packages}

# Set working directory and load relevant libraries
setwd("/Users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/")

library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggh4x)
library(clusterProfiler)
library(org.Mm.eg.db)
library(msigdbr)
library(enrichplot)
```

We have already generated the differential expression results in the last script (mPGES_11.21.23). We will now use these as an input for GSEA
```{r load in the differential expression results}

# Set the path to the differential expression files
DEG_folder <- "/users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/All_Comparions_DEGs"

# List the DEG files
DEG_files <- list.files(DEG_folder, pattern = ".csv$", full.names = FALSE)


# Initialize an empty list to hold all the DEG dataframes
DEG_results <- list()

# For loop to read in the file and store its contents in the DEG_results list
for (file in DEG_files) {
  
  # Define the full file path
  full_path <- file.path(DEG_folder, file)

  # Read in the current file
  current_file <- read.csv(full_path, header = TRUE, sep = ",")
  print(dim(current_file))
  
  # Store the data frame in the list
  DEG_results[[file]] <- current_file
  
  # Print the file name
  cat("Reading file:", file, "\n")
}
```
Now that all the files are stored in the DEG_results list, we can begin pre-processing them for GSEA analysis.
We need to order the DEGs by their log2FoldChange, create a named vector of foldChange and Entrez ID, and delete duplicated Entrez entries
```{r preprocess for GSEA}

# Initialize an empty list to store the ordered dataframes
ordered_dfs <- list()

# For loop to order dataframes by log2FoldChange
for (df in names(DEG_results)) {
 ordered <- DEG_results[[df]][order(DEG_results[[df]]$log2FoldChange, decreasing = TRUE),]
 ordered_dfs[[df]] <- ordered
}

# Initialize an empty list to hold the Entrez-named ordered foldChange vectors
entrez_foldChanges <- list()

# For each ordered dataframe, get a list of fold changes and name them by Entrez ID, remove duplicate Entrez IDs
for (df in names(ordered_dfs)) {
  # Get the fold change values and name them by Entrez ID
  ordered_genes <- ordered_dfs[[df]]$log2FoldChange
  names(ordered_genes) <- ordered_dfs[[df]]$Entrez
  
  # Identify duplicated Entrez IDs and remove them, resort in decreasing order for safety
  unique_entrez_genes <- names(ordered_genes[!duplicated(names(ordered_genes))])
  unique_genes <- ordered_genes[unique_entrez_genes]
  unique_genes <- sort(unique_genes, decreasing = TRUE)
  
  # Append Entrez-named ordered foldChange vectors to the list
  entrez_foldChanges[[df]] <- unique_genes
}
```

Let's set everything up to run GSEA. We need to create the databases for canonical pathways
```{r set up canonical pathways}

# Obtain the canonical pathways gene set for mice from Msigdbr
canonical <- msigdbr(species = "Mus musculus", category = "C2") %>%
  dplyr::select(gs_name, entrez_gene, gs_subcat) %>%
  filter(gs_subcat %in% c("CP:BIOCARTA", "CP:KEGG", "CP:PID", "CP:REACTOME", "CP:WIKIPATHWAYS"))
```

Now that we have a list of Entrez-named ordered foldChanges for each DEG file, we can begin doing the GSEA analysis
We need to iterate through each Entrez-named ordered foldChange vector and run GSEA for GO, KEGG, and canonical pathways and save the results as a .Rds object for downstream analysis. We will go this in 3 separate loops for ease of running
```{r run GSEA on each named vector for GO terms}

#GO GSEA
# Set working directory
setwd("/users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/RDS_Files/GSEA/GO/")

# For loop to run GSEA (GO) on each named vector
for (geneList in names(entrez_foldChanges)) {
  # Set seed for reproducibility
  set.seed(03061999)
  
  # Run gseGO
  gse <- gseGO(entrez_foldChanges[[geneList]],
             ont = "all",
             OrgDb = org.Mm.eg.db,
             pvalueCutoff = 0.05,
             pAdjustMethod = "BH",
             eps = 1e-300,
             verbose = TRUE,
             by = "fgsea")
  
  # Save the rds file
  saveRDS(gse, file = paste(geneList, "gseGO.rds", sep = "_"))
}
```

```{r run GSEA on each named vector for KEGG terms}

# KEGG gsea
# Set working directory
setwd("/users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/RDS_Files/GSEA/KEGG/")

# For loop to run GSEA (KEGG) on each named vector
for (geneList in names(entrez_foldChanges)) {
  # Set seed for reproducibility
  set.seed(03061999)
  
  # Run gseKEGG
  kegg <- gseKEGG(entrez_foldChanges[[geneList]],
                organism = "mmu",
                keyType = "kegg",
                pvalueCutoff = 0.05,
                pAdjustMethod = "BH",
                eps = 1e-300,
                verbose = TRUE,
                by = "fgsea")
  
  # Save the rds file
  saveRDS(kegg, file = paste(geneList, "KEGG.rds", sep = "_"))
}
```

```{r run GSEA on each named vactor for M2 Canonical Pathways}

# M2 Canonical Pathways
# Set working directory
setwd("/users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/RDS_Files/GSEA/M2_Canonical/")

# For loop to run GSEA (M2) on each named vector
for (geneList in names(entrez_foldChanges)) {
  # Set seed for reproducibility
  set.seed(03061999)
  
  # Run GSEA
  M2 <- GSEA(entrez_foldChanges[[geneList]], 
             minGSSize = 10, 
             maxGSSize = 500, 
             eps = 1e-300, 
             pvalueCutoff = 0.05, 
             pAdjustMethod = "BH", 
             by = "fgsea", 
             TERM2GENE = canonical)
   
  # Save the rds file
  saveRDS(M2, file = paste(geneList, "Canonical_Pathways.rds", sep = "_"))

}
```

