---
title: "mPGES_11.22.23 GSEA DF Results"
author: "Mike Martinez"
date: "2023-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this script is to conduct preliminary GSEA on the mPGES delta 14 KO comparisons generated in `002_mPGES_GSEA_Rds_Generator.Rmd`

```{r set working directory and load packages}

# Set working directory and load relevant libraries
setwd("/Users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/GSEA_Analysis/")

library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggh4x)
library(clusterProfiler)
library(org.Mm.eg.db)
library(msigdbr)
library(enrichplot)
```

We need to obtain the Rds objects which hold the GSEA results.
Recall in `002_mPGES_GESA_Rds_Generator.Rmd`, we generated three folders, 1 for each GSEA (GO, KEGG, and M2). Each folder should contain 8 Rds files.
We will read in the Rds files, and save the GSEA results as a table, as well as do some preliminary visualizations

```{r load in the GSEA GO Rds files}

# GO Rds folder
GO_rds_folder <- "/Users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/RDS_Files/GSEA/GO"

# List the DEG files
GO_rds_files <- list.files(GO_rds_folder, pattern = ".rds$", full.names = FALSE)

# Full path
GO_full_path <- file.path(GO_rds_folder, GO_rds_files)

# Initialize an empty list to hold all the rds files
GO_rds_list <- list()

# For loop to read in the file and store its contents in the DEG_results list
for (file in GO_full_path) {
  
  # Read in the current file
  current_file <- read_rds(file)
  
  # Store the rds object in the list
  GO_rds_list[[file]] <- current_file
  
  # Print the file name
  cat("Reading file:", file, "\n")
}

# Make the names of the list jut the basename
names(GO_rds_list) <- basename(names(GO_rds_list))
```

Now we can plot the GO enrichment dotplots
```{r plot GO dotplots}

# Set working directory
setwd("/Users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/GSEA_Analysis/GO_enrichment_dotplots/")

# Iterate through the objects and plot the enrichment dotplot
for (object in names(GO_rds_list)) {
  # Isolate just the comparison name
  object_name <- object
  result_name <- sub("_All_DEGs.csv_gseGO.rds", "", object_name)
  
  # Plot
  dotplot <- dotplot(GO_rds_list[[object]], x = "GeneRatio", color = "p.adjust", 
                           showCategory = 25, 
                           label_format = 50, 
                           split = ".sign") +
  facet_wrap(~ .sign, scales = "free_y") +
  ggtitle(result_name)
  
  # Save plot
  ggsave(paste(result_name, "GO_dotplot.png"), dotplot, width = 12, height = 10)
  
}

```


```{r save GO gsea tabular results}

# Set working directory
setwd("/Users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/GSEA_Analysis/GO_enrichment_CSV_files/")

# Iterate through the GO_rds_list and save the results as a dataframe
for (object in names(GO_rds_list)) {
  # Isolate just the comparison name
  object_name <- object
  result_name <- sub("_All_DEGs.csv_gseGO.rds", "", object_name)
  
  # Obtain the GSEA results
  result_df <- as.data.frame(setReadable(GO_rds_list[[object]], OrgDb = org.Mm.eg.db, keyType = "ENTREZID"))
  
  # Write the result as a csv file
  write.csv(result_df, file = paste(result_name, "GO_enrichment.csv", sep = "_"))
}
```
```{r load in the GSEA KEGG Rds files}

# GO Rds folder
KEGG_rds_folder <- "/Users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/RDS_Files/GSEA/KEGG"

# List the DEG files
KEGG_rds_files <- list.files(KEGG_rds_folder, pattern = ".rds$", full.names = FALSE)

# Full path
KEGG_full_path <- file.path(KEGG_rds_folder, KEGG_rds_files)

# Initialize an empty list to hold all the rds files
KEGG_rds_list <- list()

# For loop to read in the file and store its contents in the DEG_results list
for (file in KEGG_full_path) {
  
  # Read in the current file
  current_file <- read_rds(file)
  
  # Store the rds object in the list
  KEGG_rds_list[[file]] <- current_file
  
  # Print the file name
  cat("Reading file:", file, "\n")
}

# Make the names of the list jut the basename
names(KEGG_rds_list) <- basename(names(KEGG_rds_list))
```

```{r plot KEGG dotplots}

# Set working directory
setwd("/Users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/GSEA_Analysis/KEGG_enrichment_dotplots/")

# Iterate through the objects and plot the enrichment dotplot
for (object in names(KEGG_rds_list)) {
  # Isolate just the comparison name
  object_name <- object
  result_name <- sub("_All_DEGs.csv_KEGG.rds", "", object_name)
  
  # Plot
  dotplot <- dotplot(KEGG_rds_list[[object]], x = "GeneRatio", color = "p.adjust", 
                           showCategory = 25, 
                           label_format = 50, 
                           split = ".sign") +
  facet_wrap(~ .sign, scales = "free_y") +
  ggtitle(result_name)
  
  # Save plot
  ggsave(paste(result_name, "KEGG_dotplot.png"), dotplot, width = 12, height = 10)
  
}

```

```{r save KEGG gsea tabular results}

# Set working directory
setwd("/Users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/GSEA_Analysis/KEGG_enrichment_CSV_files/")

# Iterate through the GO_rds_list and save the results as a dataframe
for (object in names(KEGG_rds_list)) {
  # Isolate just the comparison name
  object_name <- object
  result_name <- sub("_All_DEGs.csv_KEGG.rds", "", object_name)
  
  # Obtain the GSEA results
  result_df <- as.data.frame(setReadable(KEGG_rds_list[[object]], OrgDb = org.Mm.eg.db, keyType = "ENTREZID"))
  
  # Write the result as a csv file
  write.csv(result_df, file = paste(result_name, "KEGG_enrichment.csv", sep = "_"))
}
```

```{r load in the GSEA Canonical Pathways (M2) Rds files}

# GO Rds folder
M2_rds_folder <- "/Users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/RDS_Files/GSEA/M2_Canonical"

# List the DEG files
M2_rds_files <- list.files(M2_rds_folder, pattern = ".rds$", full.names = FALSE)

# Full path
M2_full_path <- file.path(M2_rds_folder, M2_rds_files)

# Initialize an empty list to hold all the rds files
M2_rds_list <- list()

# For loop to read in the file and store its contents in the DEG_results list
for (file in M2_full_path) {
  
  # Read in the current file
  current_file <- read_rds(file)
  
  # Store the rds object in the list
  M2_rds_list[[file]] <- current_file
  
  # Print the file name
  cat("Reading file:", file, "\n")
}

# Make the names of the list jut the basename
names(M2_rds_list) <- basename(names(M2_rds_list))
```
```{r plot M2 dotplots}

# Set working directory
setwd("/Users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/GSEA_Analysis/M2_enrichment_dotplots/")

# Iterate through the objects and plot the enrichment dotplot
for (object in names(M2_rds_list)) {
  # Isolate just the comparison name
  object_name <- object
  result_name <- sub("_All_DEGs.csv_Canonical_Pathways.rds", "", object_name)
  
  # Plot
  dotplot <- dotplot(M2_rds_list[[object]], x = "GeneRatio", color = "p.adjust", 
                           showCategory = 25, 
                           label_format = 90, 
                           split = ".sign") +
  facet_wrap(~ .sign, scales = "free_y") +
  ggtitle(result_name)
  
  # Save plot
  ggsave(paste(result_name, "M2_dotplot.png"), dotplot, width = 20, height = 10)
  
}
```

```{r save M2 gsea tabular results}

# Set working directory
setwd("/Users/mikemartinez/Desktop/mPGES_KO/Exploratory_Analysis/GSEA_Analysis/M2_enrichment_CSV_files/")

# Iterate through the GO_rds_list and save the results as a dataframe
for (object in names(M2_rds_list)) {
  # Isolate just the comparison name
  object_name <- object
  result_name <- sub("_All_DEGs.csv_Canonical_Pathways.rds", "", object_name)
  
  # Obtain the GSEA results
  result_df <- as.data.frame(setReadable(M2_rds_list[[object]], OrgDb = org.Mm.eg.db, keyType = "ENTREZID"))
  
  # Write the result as a csv file
  write.csv(result_df, file = paste(result_name, "M2_enrichment.csv", sep = "_"))
}
```
